Hydras \& Co.

### 2023-06-21 doc/chapter-fol.tex
small changes in documentation

@@

% \subsubsection{Computing a fresh variable}

@@

The function \texttt{formula\_dec} is mainly used in
\href{../theories/html/hydras.Ackermann.PA.html}{Ackermann.PA}, in order to check whether a given formula belongs to the axioms of Peano arithmetic.
\begin{todo}
  Look for the use of \texttt{open}  (in codePA)
\end{todo}

@@

@@

\begin{todo}
  Present \texttt{substT} and \texttt{substTs} as structurally recursive functions. Then show \texttt{BadSubst.v} and motivate
 depth induction.
\end{todo}

\inputsnippets{folProp/subsTdef}

\inputsnippets{BadSubst/BadSubstFdef}

@@

For instance let $F$ be the formula 
$\exists v_2, v_1 \not= f(v_2)$. The result of the substitution 
of the term $f(v_2)$ to the (free) occurrences of $v_1$ in $F$
 \emph{is not} $\exists v_2, f(v_2) \not= f(v_2)$, but, for instance $\exists v_3, f(v_2) \not= f(v_3)$, where $v_3$ is a ``fresh'' variable.

>>>>

If we want to get the \emph{set} of variables wich occur freely 
in a formula $F$, we may use the function \texttt{List.nodup}.

@@

\begin{remark}
  The function \texttt{close} applies \texttt{freeVarF} and
  \texttt{List.nodup}  in order to add a sequence of universal
  quantifications \texttt{allH $i_1$ \dots $i_k$}, in an order determined by the implementation of these functions. It may be interesting to check whether the proof of properties of \texttt{close} depend or not from this implementation.
\end{remark}
% \subsubsection{Computing a fresh variable}

@@

The function \texttt{formula\_dec} is mainly used in
\href{../theories/html/hydras.Ackermann.PA.html}{Ackermann.PA}, in order to check whether a given formula belongs to the axioms of Peano arithmetic.

\begin{todo}
  Look for the use of \texttt{open}  (in codePA)
\end{todo}

@@

The substitution of a term to the free occurrences of a given variable $v$ is at the heart of the implementation of universal
quantifier elimination.

@@

Substitution of a term $t$ to all the occurrences of a variable $x$ in a term or a vector of terms is easy to define as a pair of mutually structurally recursive functions.

\inputsnippets{folProp/subsTdef}
\inputsnippets{FolExamples/substTExample}

Concerning formulas, it could be tempting to define substitution the same way.
\inputsnippets{BadSubst/BadSubstFdef}

Let us consider for instance the formula $F = \forall\,v_1, \,\exists\,v_2,\,v_1\not=f(v_2)$ (satisfiable if we take for instance $f$ to be interpreted as the successor function).

If we eliminate the universal quantifier by substituting in
the sub-formula $F_1=  \exists\,v_2,\,v_1\not=f(v_2)$ the free
occurrences of $v_1$ with $f(v_2)$, our naive implementation of
\texttt{substF} returns the absurd proposition $\exists\,v_2,\,f(v_2)\not=f(v_2)$. We  say that the free occurrence of $v_2$ in the term $f(v_2)$ has been \emph{captured} by the binding $\exists\,v_2,\dots$.

@@

In our example, we can obtain a correct result, if
\begin{enumerate}
\item We consider a \emph{fresh} variable, \emph{i.e.}  different from $v_1$ and $v_2$, say for instance $v_3$,
\item we substitute $v_3$ to $v_2$ in $F_1$ which results in
   $F_2= \exists\,v_3,\,v_1\not=f(v_3)$
\item we substitute the term $f(v_2)$ to $v_1$ in $F_2$, which gives us 
  $\exists\,v_3,\,f(v_2)\not=f(v_3)$.
\end{enumerate}

The notion of fresh variable is implemented through a function
\texttt{newvar (l: list nat) : nat} which returns a number which doesn't belong to $l$.

But the following attempt fails, because the renaming of a  variable in a 
sub-formula of a formula $F$ is not structurally smaller than $F$.

\inputsnippets{BadSubst/BadSubstF2}

%ICI
